Nondeterministic Bigraphical Reactive Systems for Markov Decision Processes\thanks{Supported by an EPSRC Vacation Scholarship.}

### 2018-08-16 paper.tex
PRISM section (kind of) complete

\documentclass[runningheads]{llncs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{wasysym}
\usepackage[lighttt]{lmodern}
\usepackage{siunitx}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{phaistos}
\usepackage{subcaption}
\usepackage{booktabs}

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    % outside walls
    \draw (3, 2) -- (3, 0) -- (1, 0) -- (1, 2) -- (0, 2) -- (0, 3) -- (3, 3) -- (3, 2) -- (2, 2);
    % the probabilistic wall
    \draw[red,ultra thick,dashed] (2, 2) -- (2, 1);
    % the background grid
    \draw[gray,very thin] (1, 3) -- (1, 2);
    \draw[gray,very thin] (2, 3) -- (2, 2);
    \draw[gray,very thin] (1, 1) -- (3, 1);
    \draw[gray,very thin] (2, 1) -- (2, 0);
    % the agent
    \node at (0.5, 2.3) {\PHchild};
    % the door
    \draw[dotted] (1.25,2) arc (180:270:0.5cm);
    \draw[brown,ultra thick] (1.75,1.5) -- (1.75,2) node[right,  near start]{};
    \draw (1, 2) -- (1.25, 2);
    \draw (1.75, 2) -- (2, 2);
  \end{tikzpicture}
  \caption{The full map. The agent is in the top left cell. There is a door
    connecting the middle cell of the top row with a $2 \times 2$ room below.
    The dashed wall exists with probability $1-p$.}
  \label{agent2_map}
\end{figure}

@@

The room is a $2 \times 2$ grid of \texttt{Cell}s, which is generated when the
\texttt{Agent} gets to a \texttt{Cell} with an empty \texttt{Node} (represented
by \texttt{Node\{n\}.1})\footnote{Multiple different rooms can be implemented by
having each previously empty \texttt{Node} contain a node of different control.
This way a different probability distribution (with different possible outcomes)
can be assigned to each room.}. In order for the room to be generated before any
other action can take place, the two reaction rules in Listing~\ref{agent2_room}
are put on a higher priority class. With probability $p$, the generated room has
each \texttt{Cell} linked to its two neighbours; and with probability $1-p$, the
top two \texttt{Cell}s are not linked, possibly forcing the agent to choose a
different path (see Fig.~\ref{agent2_map}). Finally, we have four actions, each
with a single reaction rule with probability $1$, corresponding to the four
possible directions of travel. The reaction rules are very similar to those of
our previous example, an example of which can be found in
Listing~\ref{agent2_north}.

\subsection{Exporting to PRISM}

% TODO: need to cite something about the syntax, logic, etc.
Now that we have two models defined in BigraphER, we will briefly show how they
can be exported and used for quantitative analysis with PRISM.

For our first model, we set the probability of finding an object to $p = 0.5$
and keep the same $2 \times 2$ grid. We ask PRISM to compute \texttt{Pmax=? [ F
  "goal" ]}, i.e., the maximum probability of reaching the goal. Recall that the
goal is to collect $n$ objects. We vary $n$ from $1$ to $3$ (as collecting $0$
objects is trivial, and collecting more than $3$ is impossible given the size of
the grid) and show the resulting probabilities in Table \ref{agent1_prism}. The
reader is welcome to check that the numbers are as expected.

@@

For the second model..

@@

\begin{itemize}
\item LIMITATIONS
  \begin{itemize}
  \item no easy way to change probabilities depending on the state
  \item each room only has a single entrance/exit (could be fixed via
    bigraphs with sharing)
  \item can't implement everything you can think of: no way to check for set
    membership, limited support for representing numbers (adding/subtracting $n$
    has to be a separate reaction rule for each $n$)
  \end{itemize}
\item EXTENSIONS
  \begin{itemize}
  \item discovery of spaces independent of room structure (no specific ideas
    yet, but could be very nice)
  \item there's no good reason why rewards are assigned to actions rather than
    reaction rules
  \item multiple agents
  \item uncertainty in actions (like in Artificial Intelligence: A Modern
    Approach)
  \end{itemize}
\item ADVANTAGES
  \begin{itemize}
  \item less code than with PRISM
  \item more visual and intuitive, represents the modelled situation in a more
    straightforward way
  \item rewards and transition rules can be generated with a higher degree of
    automation (considering a specific property of a state, rather than listing
    all states that satisfy that property)
  \item more flexible in changing the size of the grid
  \item can be easily extended to more than four directions
  \end{itemize}
\end{itemize}

>>>>

\documentclass[runningheads]{llncs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{wasysym}
\usepackage[lighttt]{lmodern}
\usepackage{siunitx}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{phaistos}
\usepackage{subcaption}
\usepackage{booktabs}

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    % outside walls
    \draw (3, 2) -- (3, 0) -- (1, 0) -- (1, 2) -- (0, 2) -- (0, 3) -- (3, 3) -- (3, 2) -- (2, 2);
    % probabilistic wall
    \draw[red,ultra thick,dashed] (2, 2) -- (2, 1);
    % background grid
    \draw[gray,very thin] (1, 3) -- (1, 2);
    \draw[gray,very thin] (2, 3) -- (2, 2);
    \draw[gray,very thin] (1, 1) -- (3, 1);
    \draw[gray,very thin] (2, 1) -- (2, 0);
    % agent
    \node at (0.5, 2.3) {\PHchild};
    % door
    \draw[dotted] (1.25,2) arc (180:270:0.5cm);
    \draw[brown,ultra thick] (1.75,1.5) -- (1.75,2) node[right,  near start]{};
    \draw (1, 2) -- (1.25, 2);
    \draw (1.75, 2) -- (2, 2);
    % target
    \node at (2.5, 1.5) {$\bigstar$};
  \end{tikzpicture}
  \caption{The full map. The agent is in the top left cell. There is a door
    connecting the middle cell of the top row with a $2 \times 2$ room below.
    The dashed wall exists with probability $p$. The star denotes the goal
    location used in Section \ref{prism_section}.}
  \label{agent2_map}
\end{figure}

@@

The room is a $2 \times 2$ grid of \texttt{Cell}s, which is generated when the
\texttt{Agent} gets to a \texttt{Cell} with an empty \texttt{Node} (represented
by \texttt{Node\{n\}.1})\footnote{Multiple different rooms can be implemented by
having each previously empty \texttt{Node} contain a node of different control.
This way a different probability distribution (with different possible outcomes)
can be assigned to each room.}. In order for the room to be generated before any
other action can take place, the two reaction rules in Listing~\ref{agent2_room}
are put on a higher priority class. With probability $1-p$, the generated room
has each \texttt{Cell} linked to its two neighbours; and with probability $p$,
the top two \texttt{Cell}s are not linked, possibly forcing the agent to choose
a different path (see Fig.~\ref{agent2_map}). Finally, we have four actions,
each with a single reaction rule with probability $1$, corresponding to the four
possible directions of travel. The reaction rules are very similar to those of
our previous example, an example of which can be found in
Listing~\ref{agent2_north}.

\subsection{Exporting to PRISM} \label{prism_section}

Now that we have two models defined in BigraphER, we will briefly show how they
can be exported and used for quantitative analysis with PRISM. More information
about the syntax and capabilities of PRISM properties can be found in the PRISM
manual\footnote{\url{http://www.prismmodelchecker.org/manual/PropertySpecification/Introduction}}.

For our first model, we set the probability of finding an object to $p = 0.5$
and keep the same $2 \times 2$ grid. We ask PRISM to compute \texttt{Pmax=? [ F
  "goal" ]}, i.e., the maximum probability of reaching the goal. Recall that the
goal is to collect $n$ objects. We vary $n$ from $1$ to $3$ (as collecting $0$
objects is trivial, and collecting more than $3$ is impossible given the size of
the grid) and show the resulting probabilities in Table \ref{agent1_prism}. The
reader is invited to check that the numbers are as expected.

@@

For the second model, we set the probability of having an extra wall in the
generated room to $p = 0.9$ and add a new atomic control called \texttt{Goal} to
denote a target location for the agent to move to. We modify the \texttt{room}
action to place the goal at the position marked by the star in
Fig.~\ref{agent2_map}. In order to count ``time'' (i.e. how many transitions are
taken), we add a predicate satisfied by every state (\texttt{Agent}) with a
reward of $1$. We also define a predicate \texttt{\textbf{big} goal =
  Cell\{x\}.(Agent | Goal | id)} that checks if the agent has reached the target
location. With these modifications in place, we can ask PRISM to compute
\texttt{Rmin=? [ F "goal" ]}, i.e., the minimum number of transitions needed to
arrive at the target location. The answer given by PRISM is \num{5.8}, and we
can check that this is indeed the correct answer: the agent takes $4$ steps
\SI{10}{\percent} of the time and $6$ steps when it needs to move around the
wall (remember to include the transition for generating the room).

@@

\begin{itemize}
\item LIMITATIONS
  \begin{itemize}
  \item no easy way to change probabilities depending on the state
  \item each room only has a single entrance/exit (could be fixed via
    bigraphs with sharing)
  \item can't implement everything you can think of: no way to check for set
    membership, limited support for representing numbers (adding/subtracting $n$
    has to be a separate reaction rule for each $n$)
  \item PRISM doesn't support importing transition rewards, positive and
    negative rewards at the same time, importing SMGs.
  \end{itemize}
\item EXTENSIONS
  \begin{itemize}
  \item discovery of spaces independent of room structure (no specific ideas
    yet, but could be very nice)
  \item there's no good reason why rewards are assigned to actions rather than
    reaction rules
  \item multiple agents
  \item uncertainty in actions (like in Artificial Intelligence: A Modern
    Approach)
  \end{itemize}
\item ADVANTAGES
  \begin{itemize}
  \item less code than with PRISM
  \item more visual and intuitive, represents the modelled situation in a more
    straightforward way
  \item rewards and transition rules can be generated with a higher degree of
    automation (considering a specific property of a state, rather than listing
    all states that satisfy that property)
  \item more flexible in changing the size of the grid
  \item can be easily extended to more than four directions
  \end{itemize}
\end{itemize}
